---
title: 'Homework 2: [YOUR NAME HERE]'
subtitle: 'Introduction to Time Series, Fall 2023'
date: 'Due Thursday September 21 at 5pm'
output:
  html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, autodep = TRUE, cache.comments = TRUE)
```

The total number of points possible for this homework is 35. The number of 
points for each question is written below, and questions marked as "bonus" are 
optional. Submit the **knitted html file** from this Rmd to Gradescope. 

If you collaborated with anybody for this homework, put their names here:

# Simple regression

1. (2 pts) 
Derive the population least squares coefficients, which solve 
\[
\min_{\beta_1, \beta_0} \, \mathbb{E} \big[ (y - \beta_0 - \beta_1 x)^2 \big],
\]
by differentiating the criterion with respect to each $\beta_j$, setting equal
to zero, and solving. Repeat the calculation but without intercept (without the
$\beta_0$ coefficient in the model). 

SOLUTION GOES HERE

2. (2 pts)
As in Q1, now derive the sample least squares coefficients, which solve
\[
\min_{\beta_1, \beta_0} \, \sum_{i=1}^n (y_i - \beta_0 - \beta_1 x_i)^2.
\]
Again, repeat the calculation but without intercept (no $\beta_0$ in the model). 

SOLUTION GOES HERE

3. (2 pts)
Prove of disprove: in the model without intercept, is the regression coefficient 
of $x$ on $y$ the inverse of that from the regression of $y$ on $x$? Answer the 
question for each of the the population and sample versions.

SOLUTION GOES HERE

4. (3 pts)
Consider the following hypothetical. Let $y$ be the height of a child and $x$ be
the height of their parent, and consider a regression of $y$ on $x$, performed 
in in a large population. Suppose that we estimate the regression coefficients 
separately for male and female parents (two separate regressions) and we find 
that the slope coefficient from the former regression $\hat\beta_1^{\text{dad}}$ 
is smaller than that from the latter $\hat\beta_1^{\text{mom}}$. Suppose however
that we find (in this same population) the sample correlation between a father's 
height and their child's height is *larger* than that between a mother's height 
and their child's height. What is a plausible explanation for what is happening 
here? 

SOLUTION GOES HERE

# Multiple regression

5. (2 pts)
In class, we claimed that the multiple regression coefficients, with respect to
responses $y_i$ and feature vectors $x_i \in \mathbb{R}^p$, $i = 1,\dots,n$, can 
be written in two ways: the first is
\[
\hat\beta = \bigg( \sum_{i=1}^n x_i x_i^T \bigg)^{-1} \sum_{i=1}^n x_i y_i.
\]
The second is 
\[
\hat\beta = (X^T X)^{-1} X^T y,
\]
where $X \in \mathbb{R}^{n \times p}$ is a feature matrix, with $i^{\text{th}}$ 
row $x_i$, and $y \in \mathbb{R}^n$ is a response vector, with $i^{\text{th}}$ 
component $y_i$. Prove that these two expressions are equivalent.

SOLUTION GOES HERE

6. (Bonus)
Derive the population and sample multiple regression coefficients by solving the
corresponding least squares problem (differentiating the criterion with respect
to each $\beta_j$, setting equal to zero, and solving). For the sample least
squares coefficient, deriving either representation in Q5 will be fine.

SOLUTION GOES HERE

# Marginal-multiple connection

7. (1 pts)
Consider the simple linear regression of a generic response $y_i$ on a constant 
predictor $x_i = 1$, $i = 1,\dots,n$, without intercept. Give the exact form of 
the sample regression coefficient.

SOLUTION GOES HERE

8. (3 pts)
Recall the connection between multiple and marginal regression coefficients, as
covered in lecture: the $j^{\text{th}}$ multiple regression coefficient can be
written in general as
\[
\hat\beta_j = \frac{(\hat{x}^{-j}_j)^T \hat{y}^{-j}}
{(\hat{x}^{-j}_j)^T \hat{x}^{-j}_j},
\]
which we interpret as the simple linear regression coefficient of $\hat{y}^{-j}$ 
on $\hat{x}^{-j}$. These are $y$ and $x_j$, respectively, after we regress out
the contributions of all other features. (See the lecture notes for the precise 
details.)

    Now note that we can treat a simple linear regression with an intercept term 
as a multiple regression with two features, with the first feature just equal to
the constant 1.  Using the above formula, and the answer from Q7, re-derive the 
expression for the slope in the simple linear model with intercept: 
\[
\hat\beta_1 = \frac{\sum_{i=1}^n (x_i - \bar{x}) (y_i - \bar{y})} 
  {\sum_{i=1}^n (x_i - \bar{x})^2}.
\]

SOLUTION GOES HERE

# Covariance calculations

9. (3 pts)
Let $x \in \mathbb{R}^n$ and $y \in \mathbb{R}^m$ be random vectors, and let 
$A \in \mathbb{R}^{k \times n}$ and $B \in \mathbb{R}^{\ell \times m}$ be fixed
matrices. Prove that
\[
\mathrm{Cov}(Ax, By) = A \mathrm{Cov}(x, y) B^T.
\]
Prove as a consequence that $\mathrm{Cov}(Ax) = A \mathrm{Cov}(x) A^T$. Hint:
you may use the rule for covariances of linear combinations (as reviewed in the 
lecture from week 2, "Measures of dependence and stationarity").

SOLUTION GOES HERE

10. (2 pts)
Suppose that $y = X \beta + \epsilon$, with $X$ and $\beta$ fixed, and where 
$\epsilon$ is a vector with white noise entries, with variance $\sigma^2$. Use 
the rule in Q9 to prove that for the sample least squares coefficients, namely, 
$\hat\beta = (X^T X)^{-1} X^T y$, it holds that
\[
\mathrm{Cov}(\hat\beta) = \sigma^2 (X^T X)^{-1}.
\]

SOLUTION GOES HERE

11. (4 pts)
An equivalent way to state the Gauss-Markov theorem is as follows. Under the 
model from Q10, if $\tilde\beta$ is any other unbiased linear estimator of
$\beta$ (where linearity means that $\tilde\beta = My$ for a fixed matrix $M$)
then 
\[
\mathrm{Cov}(\hat\beta) \lesssim \mathrm{Cov}(\tilde\beta)
\]
where $\lesssim$ means less than or equal to in the *PSD (positive semidefinite)
ordering*. Precisely, $A \lesssim B$ if and only if $B-A$ is a PSD matrix, which
recall, means $z^T (B-A) z \geq 0$ for all vectors $z$. Prove that this is 
indeed equivalent to the statement of the Gauss-Markov theorem given in lecture. 

SOLUTION GOES HERE

# Cross-validation

12. (3 pts)
Recall the R code from lecture that performs time series cross-validation to 
evaluate the mean absolute error (MAE) of predictions from the linear regression
of cardiovascular mortality on 4-week lagged particulate levels. (The version 
where the regression is trained on all-past). Adapt this to evaluate the MAE of
predictions from a multiple linear regression of cardiovascular mortality on 
4-week lagged particulate levels and 4-week lagged temperature (2 features). 
Plot the predictions, and print the MAE on the plot, following the lecture code.

    Additionally (all drawn on the same figure), plot the fitted values on the 
training set. By the training set here, we mean what is also called the "burn-in
set" in the lecture notes, and indexed by times 1 through `t0` in the code. The 
fitted values should come from the initial regression model that is fit to the 
burn-in set. Print the MAE from the fitted values the training set somewhere on 
the plot (and label this as "Training MAE" to clearly differentiate it).

```{r}
# CODE GOES HERE
```

13. (2 pts)
Repeat the same exercise as in Q12 but now with multiple lags per variable: use
lags 4, 8, 12 for each of particulate level and temperature (thus 6 features in 
total). Did the training MAE go down? Did the cross-validated MAE go down? 
Discuss. Hint: you may find it useful to know that `lm()` can take a predictor
matrix, as in `lm(y ~ x)` where `x` is a matrix; in this problem, you can form 
the predictor matrix by calling `cbind()` on the lagged feature vectors.

```{r}
# CODE GOES HERE
```

14. (2 pts)
Repeat once more the same exercise as in the last question but but now with many
lags per variable: use lags 4, 5, ..., through 50 for each of particulate level 
and temperature (thus 47 x 2 = 94 features in total). Did the training MAE go 
down? Did the cross-validated MAE go down? Are you surprised? Discuss.

```{r}
# CODE GOES HERE
```

# More features, the merrier?

15. (2 pts)
Let $y_i$ be an arbitrary response, and $x_i \in \mathbb{R}^p$ be an arbitrary 
feature vector, for $i = 1,\dots,n$. Let 
\[
\tilde{x}_i = (x_{i1}, \dots, x_{ip}, \tilde{x}_{i,p+1}),\quad i = 1,\dots,n
\]
be the result of appending one more feature. Let $\hat{y}_i$ denote the fitted 
values from the regression of $y_i$ on $x_i$, and let $\tilde{y}_i$ denote the 
fitted values from the regression of $y_i$ on $\tilde{x}_i$. Prove that 
\[
\sum_{i=1}^n (y_i - \tilde{y}_i)^2 \leq \sum_{i=1}^n (y_i - \hat{y}_i)^2.
\]
In other words, *the training MSE will never get worse as we add features* to a 
given sample regression problem.

SOLUTION GOES HERE

16. (2 pts) 
How many linearly independent features do we need (how large should $p$ be) in 
order to achieve a perfect training accuracy, i.e., training MSE of zero? Why?

SOLUTION GOES HERE

17. (Bonus)
Implement an example in R in order to verify your answer to Q16 empirically. 
Extra bonus points if you do it on the cardiovascular mortality data, using
enough lagged features. You should be able to plot the fitted values from the 
training set and see that they match the observations perfectly (and the CV
predictions should look super wild).

```{r}
# CODE GOES HERE
```
