---
title: 'Homework 3: [YOUR NAME HERE]'
subtitle: 'Introduction to Time Series, Fall 2023'
date: 'Due Thursday October 5 at 5pm'
output:
  html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, autodep = TRUE, cache.comments = TRUE)
```

The total number of points possible for this homework is 33. The number of 
points for each question is written below, and questions marked as "bonus" are 
optional. Submit the **knitted html file** from this Rmd to Gradescope. 

If you collaborated with anybody for this homework, put their names here:

# Regression troubles

1. (5 pts)
Suppose that $y \in \mathbb{R}^n$ is a response vector and $X \in \mathbb{R}^{n 
\times p}$ is a predictor matrix, with $p > n$. Prove that there is at least 
one $\eta \not= 0$ (not equal to the zero vector) that is in $\mathrm{null}(X)$,
the null space of $X$. Prove that if $\tilde\beta$ is a least squares solution 
in the regression of $y$ on $X$, then any vector of the form
\[
\hat\beta = \tilde\beta + \eta, \quad \text{for $\eta \in \mathrm{null}(X)$}
\]
is also a solution.

SOLUTION GOES HERE

2. (6 pts)
With $X, y$ as in Q1, suppose that $\tilde\beta$ is a least squares solution 
with $\tilde\beta_j > 0$, and suppose that $\mathrm{null}(X) \not\perp e_j$, 
where $e_j$ is the $j^{\text{th}}$ standard basis vector (i.e., $e_j$ is a 
vector with all 0s except for a 1 in the $j^{\text{th}}$ component), and recall 
we write $S \perp v$ for a set $S$ and vector $v$ provided $u^T v = 0$ for all
$u \in S$. Prove that there exists another least squares solution $\hat\beta$ 
such that $\hat\beta_j < 0$.

SOLUTION GOES HERE

# Ridge and lasso

3. (8 pts)
On the cardiovascular mortality regression data, form lagged features from the
particulate matter and temperature variables, using lags 4, 8, ..., 40 from 
each. Using the `glmnet` package, fit a ridge regression and lasso regression
(two separate models), each over a grid of tuning parameter values $\lambda$ 
chosen by the `glmnet()` function, with cardiovascular mortality as the response
and all the lagged features as predictors (you should have 20 in total: 10 from
particulate matter, and 10 from temperature). However, make sure you do this in
a split-sample setup for validation, as follows, for each of ridge and lasso:

- fit the `glmnet` object on the *first half of the time series*; 
- make predictions on the *second half of the time series*, for each $\lambda$;
- record the MAE of the predictions on the second half, for each $\lambda$;
- choose and report the value of $\lambda$ with the lowest MAE;
- plot the cardiovascular mortality time series, along with the predictions on 
  the second half, and print the MAE and the selected value of $\lambda$ on the 
  plot.

    You can build off the code given in the regularization lecture (weeks 5-6: 
"Regularization and smoothing") for fitting the ridge and lasso models, and the 
regression lecture (weeks 3-4: "Regression and prediction") for the split-sample
validation. *Note carefully that the lecture code includes lag 0, and here we do
not, so that we can make ex-ante 4-week ahead forecasts!*

```{r}
# CODE GOES HERE
```

4. (1 pts)
Which lagged features were present in the MAE-optimal lasso model, selected by
split-sample validation, in Q3?

```{r}
# CODE GOES HERE
```

5. (8 pts)
Repeat Q3, except implement time series cross-validation instead of split-sample 
validation. You should begin time series cross-validation on the second half of
the time series, treating the first half as a burn-in set. Also, be sure to fit 
each ridge or lasso model using a trailing window of 200 time points (not all 
past). 

    Warning: doing time series cross-validation properly will require us to pay 
attention to the following. The `glmnet()` function chooses a sequence of tuning 
parameter values $\lambda$ based on the passed feature matrix `x` and response 
vector `y` (its first two arguments). However, in time series cross-validation, 
this will change at each time point. So if you just call `glmnet()` naively, 
then you will not have a consistent $\lambda$ sequence over which to calculate 
MAE and perform tuning.

    You can circumvent this issue by defining your own $\lambda$ sequence ahead
of time, and forcing `glmnet()` to use it by passing it through its `lambda` 
argument. Indeed, the best thing to do here is just to use the `lambda` sequence
that `glmnet()` itself derived for the ridge and lasso models fit to the first
half of the time series, which you already have from Q3. Do this, and then just
as in Q3, produce a plot of the cardiovascular mortality time series, along with
the predictions from time series CV on the second half, and print the MAE and 
the selected value of $\lambda$ on the plot.

    You can build off the code given in the regression lecture for time series
cross-validation (or the code you wrote in Homework 2 to implement time series
cross-validation).

```{r}
# CODE GOES HERE
```

6. (Bonus) 
Report which lagged features were most frequently selected by the lasso. Because
the lasso models will be refit at each time point (in the second half of the
data set), you will have to additionally store the lasso solutions along your 
time series CV pass. Then, look back at the solutions that correspond to the 
MAE-optimal $\lambda$ value, and choose some way of summarizing which of its 
components were consistently large in magnitude over time.

```{r}
# CODE GOES HERE
```

# HP filter

7. (5 pts)
Recall in lecture we saw the HP filter could be written explicitly as
\[
\hat\theta = \underbrace{(I + \lambda D^T D)^{-1}}_K \, y,
\]
where $D \in \mathbb{R}^{(n-2) \times n}$ is the second difference matrix on $n$
points. In other words, defining $K \in \mathbb{R}^{n \times n}$ as above,
\[
\hat\theta_i = \sum_{j=1}^n K_{ij} y_j, \quad i = 1,\dots,n.
\]
Compute the matrix $K$ empirically for a problem of size $n=100$, and setting 
the tuning parameter to be $\lambda=100$; inspect three of its rows, at indices
$i = 25, 50, 75$. For each $i$, plot the $i^{\text{th}}$ row as a curve over the 
underlying position $1,\dots,n$; that is, plot the x-y pairs
\[
(x_j, y_j) = (j, K_{ij}), \quad j = 1,\dots,n
\]
as a curve. Overlay the curves for all three rows on the same plot, each in a 
different color. What do these curves look like to you? Use the plot to argue 
that the HP filter acts like a kernel smoother.

```{r}
# CODE GOES HERE
```

8. (Bonus) 
Do a literature search to find theory on the *asymptotically equivalent kernel* 
for the HP filter. This should have a closed-form. Plot this and comment on 
whether or not your empirical results adhere to what is known asymptotically.

```{r}
# CODE GOES HERE
```

# Trend filter

9. (Bonus)
Implement 5-fold cross-validation in order to tune $\lambda$ in trend filtering
applied to the Boston marathon men's data set. Recall the description of how to
set folds in a special "structured" way, for tuning smoothers, given near the 
end of the lecture notes (weeks 5-6: "Regularization and smoothing"). The code
below shows how to run trend filtering and derive estimates at the held-out
points for one fold. You can build off this code for your solution. You will 
need to install the `glmgen` package from GitHub, which you can do using the 
code that has been commented out. 

    Important note: just like `glmnet()`, the `trendfilter()`  function (in the
`glmnet` package) computes its own `lambda` sequence. So you will need to define 
an initial `lambda` sequence to pass to each subsequent call to `trendfilter()`,
so that you can have a consistent grid of tuning parameter values over which to
perform cross-validation. We do this below by using the `lambda` sequence that
`trendfilter()` itself derived when the trend filtering model is fit on the full
data set.

    After implementing cross-validation, compute and report the $\lambda$ value
with the smallest cross-validated MAE. Then, lastly, plot the solution at this
value of $\lambda$ when the model is fit to the full data set (this is already
available in the `tf` object below.)

```{r}
# devtools::install_github("glmgen/glmgen", subdir = "R_pkg/glmgen")
library(glmgen)
library(fpp3)

boston = boston_marathon |>
  filter(Year >= 1924) |>
  filter(Event == "Men's open division") |>
  mutate(Minutes = as.numeric(Time)/60) |>
  select(Year, Minutes)

# Fit trend filtering on the entire data in order to grab the lambda sequence
tf = trendfilter(x = boston$Year, y = boston$Minutes, k = 1)
lambda = tf$lambda

n = nrow(boston)       # Number of points
k = 5                  # Number of folds
inds = rep_len(1:k, n) # Folds indices

# Fit trend filtering on all points but those in first fold. We are forcing it
# to use the lambda sequence that we saved above
tf_subset = trendfilter(x = boston$Year[inds != 1], 
                        y = boston$Minutes[inds != 1], 
                        k = 1, lambda = lambda)

# Compute the predictions on the points in the first fold. Plot the predictions
# (as a sanity check) at a particular value of lambda in the middle of the grid
yhat = predict(tf_subset, x.new = boston$Year[inds == 1])
plot(boston$Year, boston$Minutes, col = 8)
points(boston$Year[inds == 1], yhat[, 25], col = 2, pch = 19)
```
